---
title: "homework 6"
output: 
  html_document: 
    highlight: monochrome
    keep_md: yes
    number_sections: yes
    theme: journal
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
setwd("C:\\Users\\Steven\\Google Drive\\1. MOT\\1) Fall 2016\\1. BA\\Assignments\\Homework 6")
 
sp <- read.csv("sp.csv", stringsAsFactors=FALSE, header=TRUE)
```

```{r}
sp$Date <- as.Date(sp$Date, "%d-%B-%y")
```


#sort by date, starting with the oldest dates
```{r}
sp <- sp[order(as.Date(sp$Date, format="%Y/%m/%d", decreasing=FALSE)),]
```



#sort in decreasing fashion using reverse function (decreasing=TRUE does not work for dates)

```{r}
sp_descending <- sp[rev(order(as.Date(sp$Date))),]

```


#read in gold 
```{r}
gold <- read.csv("gold2016.csv", stringsAsFactors=FALSE, header=TRUE)
gold$Date <- as.Date(gold$Date, "%d-%B-%y")
gold <- gold[order(as.Date(gold$Date, format="%y/%m/%d")),]
```




#create monthly intervals by just third day 

```{r}
gold['day']<- strftime(gold$Date, "%d")
```

#create snapshot monthly price of Gold on third of month 
```{r}
gold_month <- subset(gold, gold$day == "03")
gold_month['gold_month_price'] <- gold_month$Price
gold_month <- subset(gold_month, select=c("gold_month_price"))

```


#calculate average monthly price of Gold 
```{r}
gold['day_year'] <- strftime(gold$Date, "%m/%y")
gold['avg_price'] <- 0
for (i in 1:nrow(gold)) {
  submonth <- subset(gold, day_year == day_year[i])
  gold$avg_price[i] <- (sum(submonth$Price))/length(submonth$Price)
}
```





#create training and test data for to gague forecast accuracy 
```{r}
gold_train <- subset(gold, Date > as.Date("2016-01-01"))
gold_test <- subset(gold, Date < as.Date("2016-01-01"))
```




#read in oil
```{r}
oil <- read.csv("oil2016.csv", stringsAsFactors=FALSE, header=TRUE)
oil$Date <- as.Date(oil$Date, "%d-%B-%y")
oil <- oil[order(as.Date(oil$Date, format="%y/%m/%d")),]
```



#oil monthly 
```{r}
oil['day_year'] <- strftime(oil$Date, "%m/%y")
oil['avg_price'] <- 0
for (i in 1:nrow(oil)) {
  submonth <- subset(oil, day_year == day_year[i])
  oil$avg_price[i] <- (sum(submonth$Price))/length(submonth$Price)
}

```

#s&p500 monthly 
```{r}
sp['day_year'] <- strftime(sp$Date, "%m/%y")
sp['avg_price'] <- 0
for (i in 1:nrow(sp)) {
  submonth <- subset(sp, day_year == day_year[i])
  sp$avg_price[i] <- (sum(submonth$Price))/length(submonth$Price)
}
```



#create test/train data for oil 
```{r}
oil_train <- subset(oil, Date > as.Date("2016-01-01") )
oil_test <- subset(oil, Date < as.Date("2016-01-01"))
```




#merge
```{r}
oil['oil'] <- oil$Price 
sp['sp'] <- sp$Price
gold['gold'] <- gold$Price
gold['gold_avg'] <-gold$avg_price

timeseries <- merge(oil, sp, by="Date")
timeseries <- merge(timeseries, gold, by="Date")
timeseries <- subset(timeseries, select=c("Date", "oil", "sp", "gold", "gold_avg"))
```



#unique TS for monthly values  
```{r}
timeseries_month <- unique(timeseries$gold_avg, incomparables=FALSE)

timeseries_month_oil <- unique(oil$avg_price, incomparables=FALSE)

timeseries_month_sp <- unique(sp$avg_price, incomparables=FALSE)
```



#Create time series
```{r}
timeseries_all <- subset(timeseries, select=c("oil", "sp", "gold"))
ts.all <- ts(data=timeseries_all)

ts.oil<-ts(timeseries_all$oil, start=c(2010))
ts.oil_avg <- ts(timeseries_month_oil, start=c(2010), frequency=12)

ts.gold <- ts(timeseries_all$gold, start=c(2010), frequency=5)

```


#Gold Time Series (Daily)

```{r, echo=FALSE}
plot.ts(ts.gold)
```

```{r}
ts.gold_avg <- ts(timeseries_month, start=c(2010), frequency=12)

ts.sp <- ts(timeseries_all$sp)
ts.sp_avg <- ts(timeseries_month_sp,  start=c(2010), frequency=12 )
```



#combine oil and gold monthly 

```{r}
ts.combo <- cbind(ts.oil_avg, ts.gold_avg)
```

```{r, echo=FALSE}
plot.ts(ts.combo)
```
#Oil and Gold Monthly Correlation  
```{r, echo=FALSE}

par(mfrow=c(2, 1))

plot(ts.gold_avg)
plot(ts.oil_avg)
```

```{r}
 decompose 
ts.gold_avg.d <- decompose(ts.gold_avg)
```

#Gold Monthly Average Time Series - Decomposed 

```{r, echo=FALSE}
plot(ts.gold_avg.d)
```

```{r}
ts.oil_avg.d <- decompose(ts.oil_avg)
```

#Oil Monthly Average Time Series - Decomposed 

```{r, echo=FALSE}
plot(ts.oil_avg.d)

plot(ts.oil)
```

```{r}
ts.gold.1.d <- decompose(ts.gold_avg)

```

#Gold Monthly Average Time Series 

```{r, echo=FALSE}
plot(ts.gold.1.d)
```

```{r}
ts.sp_avg.d <- decompose(ts.sp_avg)
```

#SP 500 Monthly Average Time Series 

```{r, echo=FALSE}
plot(ts.sp_avg.d)
```


#Gold forecast with HoltWinters 

```{r}
library(forecast)

gold.holt <- HoltWinters(ts.gold_avg, gamma=FALSE)

gold.holt_fore <- forecast.HoltWinters(gold.holt, h=12)
```

#Gold Forecast (Holt Winters)

```{r, echo=FALSE}
plot(gold.holt)

plot(gold.holt_fore)
```


#Oil Forecast (Holt Winters)

```{r}
oil.holt <- HoltWinters(ts.oil_avg, gamma=FALSE)
oil.holt_fore <- forecast.HoltWinters(oil.holt, h=12)

```

#Oil Forecast (Holt Winters)

```{r, echo=FALSE}

plot(oil.holt_fore)
plot(oil.holt)
```

#S&P Forecast (Holt Winters)

```{r}
sp.holt <- HoltWinters(ts.sp_avg, gamma=FALSE)
sp.holt_fore <- forecast.HoltWinters(sp.holt, h=12)

```

#S&P Forecast (Holt Winters)

```{r, echo=FALSE}
plot(sp.holt_fore)
plot(sp.holt)
```
 

#gold seasonal (HoltWinters forecast)

```{r}
gold.holt.seasonal <- HoltWinters(ts.gold_avg, gamma=TRUE)
gold.holt.seasonal <- forecast.HoltWinters(gold.holt.seasonal, h=12)
```
```{r, echo=FALSE}
plot(gold.holt.seasonal)
```

#oil seasonal (HoltWinters forecast)

```{r}
oil.holt.seasonal <- HoltWinters(ts.gold_avg, gamma=TRUE)
oil.holt.seasonal.fore <- forecast.HoltWinters(oil.holt.seasonal)

sp.holt.seasonal <- HoltWinters(ts.sp_avg, gamma=TRUE)
sp.holt.seasonal.fore <- forecast.HoltWinters(sp.holt.seasonal)


```{r, echo=FALSE}
 
plot(oil.holt.seasonal.fore)
plot(sp.holt.seasonal.fore)
```

#test data for stationarity 

```{r}
require(tseries)
adf.test(diff(diff(log(ts.gold_avg))))
adf.test(diff(diff(log(ts.oil_avg))))
adf.test(diff(diff(log(ts.oil))))
```

#create standard ARIMA model with no differencing to see what changes are neccesary

```{r}
library(forecast)  
library(stats)
 
fit_1<- arima(ts.gold_avg, c(0,0,0))
fit_2<- arima(ts.oil_avg, c(0,0,0))
fit_3 <- arima(ts.sp_avg, c(0,0,0))

```

#ARIMA model for Gold (no differencing)

```{r, echo=FALSE}

tsdiag(fit_1)
```

#ARIMA model for Oil (no differencing)

```{r, echo=FALSE}

tsdiag(fit_2)
```

#ARIMA model for S&P500 (no differencing)

```{r, echo=FALSE}

tsdiag(fit_3)
```


#get lazy and have R create fit for you 

```{r}


fit_gold<- auto.arima(ts.gold_avg)
fit_gold <- arima(ts.gold_avg, c(1, 1, 1))

fit_oil <- auto.arima(ts.oil_avg)


fit_oil <- arima(ts.oil_avg, c(0,1, 1))


fit_sp <- auto.arima(ts.sp_avg)
 

fit_sp <- arima(ts.sp_avg, c(0,1,0))
```


#gold ARIMA 

```{r, echo=FALSE}
tsdiag(fit_gold)
```


#oil ARIMA 

```{r, echo=FALSE}
tsdiag(fit_oil)

```

#sp ARIMA

```{r, echo=FALSE}
tsdiag(fit_sp)
```


 



#create forecast 

```{r}

fit_gold <- forecast.Arima(fit_gold, h=12)
plot(fit_gold)
summary(fit_gold)

```

